Notes on generating initial models for mapping into {\tt MAESTRO}

### 2008-08-19 docs/initial_models/paper.tex
minor edits

The initial data needs be generated at the same resolution as the base state for single-level problems and at the same resolution of the finest resolution base state for multi-level plane-parallel problems.  For spherical problems, we force $\Delta r = \Delta x$.
\section{Creating the Base State and Full State}
Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.  At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.  Here are the steps:
\begin{enumerate}
\item Call the EOS to obtain a 1D $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.  Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\init} = p^{\init}(\rho^{\init},T^{\init},X^{\init})$ at the end of the previous fortran subroutine.
\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Set $\rho_0 = \overline\rho$.
\item Compute $p_0$ using equation (\ref{HSE Discretization}).
\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
\end{enumerate}
Now $\rho_0 = {\rm Avg}(\rho)$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
\end{document}

>>>>

The initial data needs be generated at the same resolution as the base state for single-level problems and at the same resolution of the finest resolution base state for multi-level plane-parallel problems.  For spherical problems, we force $\Delta r = \Delta x$.
\section{Creating the Base State and Full State}
Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.  We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.  Here are the steps:
\begin{enumerate}
\item Call the EOS to obtain a 1D $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.  Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\init} = p^{\init}(\rho^{\init},T^{\init},X^{\init})$ at the end of the previous fortran subroutine.
\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Set $\rho_0 = \overline\rho$.
\item Compute $p_0$ using equation (\ref{HSE Discretization}).
\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
\end{enumerate}
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
\end{document}
